---
title: "Chapter 5 Excercises"
author: "Jonathan Bryan"
date: "May 19, 2018"
output: pdf_document
---

### 1. Using basic statistical properties of the variance, as well as singlevariable calculus, derive (5.6). In other words, prove that $\alpha$ given by (5.6) does indeed minimize $Var(\alpha X + (1 - \alpha)Y )$.
$$
\begin{aligned}
\underset{\alpha}{\operatorname{argmin}}\; Var(\alpha X + (1 - \alpha)Y)&\\
Var(\alpha X + (1 - \alpha)Y ) &= \alpha^2 Var(X) + (1 - \alpha)^2Var(Y) + 2\alpha(1-\alpha)Cov(X,Y))\\
Var(\alpha X + (1 - \alpha)Y ) &= \alpha^2 \sigma_X^2 + (1 - 2\alpha + \alpha^2)\sigma_Y^2 + (2\alpha-2\alpha^2)\sigma_{XY}^2\\
\frac{d}{d\alpha} Var(\alpha X + (1 - \alpha)Y ) &= 2\alpha \sigma_X^2 + (-2 + 2\alpha)\sigma_Y^2 + (2-4\alpha)\sigma_{XY}^2\\
\frac{d}{d\alpha} Var(\alpha X + (1 - \alpha)Y ) &= 2\alpha \sigma_X^2 - 2\sigma_Y^2 + 2\alpha\sigma_Y^2 + 2\sigma_{XY}^2-4\alpha\sigma_{XY}^2\\
2\alpha \sigma_X^2 - 2\sigma_Y^2 + 2\alpha\sigma_Y^2 + 2\sigma_{XY}^2-4\alpha\sigma_{XY}^2 &= 0\\
2\alpha \sigma_X^2 + 2\alpha\sigma_Y^2 -4\alpha\sigma_{XY}^2 &= 2\sigma_Y^2 - 2\sigma_{XY}^2\\
2\alpha (\sigma_X^2 + \sigma_Y^2 -2\sigma_{XY}^2) &= 2\sigma_Y^2 - 2\sigma_{XY}^2\\
\alpha&= \frac{\sigma_Y^2 - \sigma_{XY}^2}{\sigma_X^2 + \sigma_Y^2 -2\sigma_{XY}^2}\\
\end{aligned}
$$

### 2. We will now derive the probability that a given observation is part of a bootstrap sample. Suppose that we obtain a bootstrap sample from a set of n observations.

#### (a) What is the probability that the first bootstrap observation is not the jth observation from the original sample? Justify your answer.
$$Pr(B^1 \neq j) = \frac{n-1}{n}$$

#### (b) What is the probability that the second bootstrap observation is not the jth observation from the original sample?
$$Pr(B^2 \neq j) = \frac{n-1}{n}$$

#### (c) Argue that the probability that the jth observation is not in the bootstrap sample is $(1 - 1/n)/^n$.
$$Pr(B^1 \neq j,B^2 \neq j,...,B^n \neq j ) = \prod_{i = 1}^n\frac{n-1}{n} = \frac{n-1}{n}^n$$

#### (d) When n = 5, what is the probability that the jth observation is in the bootstrap sample?
$$Pr(j \in \bf{B} | n = 5) = 1 - \binom{5}{0} (1/5)^0 (4/5)^5 = 1 - .328 = .672$$

#### (e) When n = 100, what is the probability that the jth observation is in the bootstrap sample?
$$Pr(j \in \bf{B} | n = 100) = 1 -  (99/100)^{100} = 1 - .366 = .634$$

#### (e) When n = 10, 000, what is the probability that the jth observation is in the bootstrap sample?
$$Pr(j \in \bf{B} | n = 10000) = 1 -  (9999/10000)^{10000} = 1 - .367 = .632$$

#### (g) Create a plot that displays, for each integer value of n from 1 to 100, 000, the probability that the jth observation is in the bootstrap sample. Comment on what you observe.
```{r}
x = seq(1,100000)
y = 1 - ((x - 1)/x)^x
plot(x,y, main = "Probability of jth Observation in Bootstrap",
          xlab = "n",
          ylab = expression(1 - ((x - 1)/x)^x))
```

The probability that the jth observation is in the bootstrap approaches $\approx \frac{2}{3}$ as n increaes. 

#### (h) We will now investigate numerically the probability that a bootstrap sample of size n = 100 contains the jth observation. Here j = 4. We repeatedly create bootstrap samples, and each time we record whether or not the fourth observation is contained in the bootstrap sample.
```{r}
store = rep(NA, 10000)
for(i in 1:10000){
  store[i] = sum(sample(1:100, rep=TRUE) ==4) > 0
}
mean(store)
```
On average the probability that the jth observation is in the bootstrap sample is .64,

### 3. We now review k-fold cross-validation.

#### (a) Explain how k-fold cross-validation is implemented.
A integer k is selected and a $\frac{k-1}{k}$ proportion of the data is randomly selected as the training set while the $\frac{1}{k}$ rest of the data serves as the holdout set. The model is trained on the training set and validated using the holdout set k times. An estimate of the test error is obtained by averaging the total error (e.g. RMSE, classification error) of each model.

#### (b) What are the advantages and disadvantages of k-fold cross-validation relative to:

##### i. The validation set approach?
k-fold cross-validation will typically result in a better estimate of the test error rate (especially if k is between 5 or 10) by reducing the bias and variance of the estimate as more data is used to train the model. However, k-fold cross-validation requires possibly fine tuning the optimal k to derive the best test error estimated, while the validation approach always divides the dataset into two parts and is thus computationally faster.

##### ii. LOOCV?
While LOOCV has the lowest bias of the cross-validation estimators of test error rates, the LOOCV test error estimate will vary more with different datasets of size $n$ from the same population due to using nearly the same training data for each of the $n$ models and thus having outputs that are highly correlated. k-fold cross-validation is usually preferred as it has a better balance of bias and variance performance. LOOCV is also computationally more expensive when $k < n$. 

### 4. Suppose that we use some statistical learning method to make a prediction for the response Y for a particular value of the predictor X. Carefully describe how we might estimate the standard deviation of our prediction.
We could repeat the process of fitting the statistical model by drawing a random observation and associated response from the data $(X^*,Y^*)$ $n$ number of times. Each fitted model, could then use the particular value of $X^p$ to to predict $\hat{Y^p}$. We would collect $n$ prediction estimates using $X^p$ and calculated the standard deviation using $\sqrt{\{\frac{1}{n-1}\Sigma_{i=1}^n(\hat{Y^p} - \bar{\hat{Y^p}})}$. 

### 5. In Chapter 4, we used logistic regression to predict the probability of default using income and balance on the Default data set. We will now estimate the test error of this logistic regression model using the validation set approach. Do not forget to set a random seed before beginning your analysis.

#### (a) Fit a logistic regression model that uses income and balance to predict default.
```{r, message=FALSE}
library(ISLR)
set.seed(1)
lr = glm(as.numeric(default)-1 ~ income + balance, family = "binomial", data = Default)
stargazer::stargazer(lr)
```

#### (b) Using the validation set approach, estimate the test error of this model. In order to do this, you must perform the following steps:

##### i. Split the sample set into a training set and a validation set.
```{r}
set.seed(1)
train = sample(seq(1,nrow(Default),by=1),size = nrow(Default)/2 , replace=FALSE)
default.train = Default[train,]
default.validation = Default[-train,]
```

##### ii. Fit a multiple logistic regression model using only the training observations.
```{r}
lr2 = glm(as.numeric(default) - 1 ~ income + balance, family = "binomial", data = default.train)
```

##### iii. Obtain a prediction of default status for each individual in the validation set by computing the posterior probability of default for that individual, and classifying the individual to the default category if the posterior probability is greater than 0.5.
```{r}
lr2.pred = predict(lr2, newdata = default.validation, type = "response")
lr2.pred[lr2.pred > 0.5] = 1
lr2.pred[lr2.pred <= 0.5] = 0
```

##### iv. Compute the validation set error, which is the fraction of the observations in the validation set that are misclassified.
```{r}
sum(lr2.pred != as.numeric(default.validation$default)-1)/nrow(default.validation)
```
The validation set error rate is 2.86%.

#### (c) Repeat the process in (b) three times, using three different splits of the observations into a training set and a validation set. Comment on the results obtained.
```{r}
set.seed(2)
class.error = rep(NA,3)
for ( i in 1:3){
  train = sample( seq(1, nrow(Default), by = 1), size = nrow(Default)/2, replace = FALSE)
  default.train = Default[train,]
  default.validation = Default[-train,]
  lr3 = glm(as.numeric(default) - 1 ~ income + balance, family = "binomial", data = default.train)
  lr3.pred = predict(lr3, newdata = default.validation, type = "response")
  lr3.pred[lr3.pred > 0.5] = 1
  lr3.pred[lr3.pred <= 0.5] = 0
  class.error[i] = sum(lr3.pred != as.numeric(default.validation$default) - 1)/nrow(default.validation)
}
range(class.error)
mean(class.error)
```

The validation set misclassification rate ranged from 2.58% to 2.78% and the mean was 2.71%.

#### (d) Now consider a logistic regression model that predicts the probability of default using income, balance, and a dummy variable for student. Estimate the test error for this model using the validation set approach. Comment on whether or not including a dummy variable for student leads to a reduction in the test error rate.
```{r}
set.seed(3)
train = sample( seq(1, nrow(Default), by = 1), size = nrow(Default)/2, replace = FALSE)
default.train = Default[train,]
default.validation = Default[-train,]
lr4 = glm(as.numeric(default) - 1 ~ income + balance + as.factor(student), family = "binomial", data = default.train)
lr4.pred = predict(lr4, newdata = default.validation, type = "response")
lr4.pred[lr4.pred > 0.5] = 1
lr4.pred[lr4.pred <= 0.5] = 0
sum(lr4.pred != as.numeric(default.validation$default) - 1)/nrow(default.validation)
```

Using a student dummy variable lowered the validation set misclassification rate to 2.48%.

### 6. We continue to consider the use of a logistic regression model to predict the probability of default using income and balance on the Default data set. In particular, we will now compute estimates for the standard errors of the income and balance logistic regression coefficients in two different ways: (1) using the bootstrap, and (2) using the standard formula for computing the standard errors in the glm() function. Do not forget to set a random seed before beginning your analysis.

