---
title: "Chapter-3-Excercises"
author: "Jonathan Bryan"
date: "July 26, 2017"
output:
  pdf_document: default
  html_document: default
---

### 1.Describe the null hypotheses to which the p-values given in Table 3.4 correspond. Explain what conclusions you can draw based on these p-values. Your explanation should be phrased in terms of sales, TV, radio, and newspaper, rather than in terms of the coefficients of the linear model.
The null hypothesis tested for the coefficients in the linear regression model asserts that there is no relationship between the predictor and the response variable, which mathematically means the coefficient is zero. At an alpha level of 0.05 we can say that both TV and radio advertising spending are significant predictors for the sales response variable. More specifically, the model says that every dollar towards TV advertising is associated with an additional \$46 in sales and every  dollar towards radio advertising is associated with \$189 dollars in sales.

### 2. Carefully explain the differences between the KNN classifier and KNN regression methods
Both KNN classifiers and regressions are non-parametric models. Both methods use a distance metric to calculate a pre-scpecified K number of data points nearest neighbors. KNN classifiers then use a voting method to predict a class based upon the labeled classes of the neighbors. KNN regressions average the numerical values of the neighbors to generate an estimate.

### 3. Suppose we have a data set with five predictors, X1 = GPA, X2 = IQ, X3 = Gender (1 for Female and 0 for Male), X4 = Interaction between GPA and IQ, and X5 = Interaction between GPA and Gender. The response is starting salary after graduation (in thousands of dollars). Suppose we use least squares to fit the model, and get $\hat{\beta_0}$ = 50, $\hat{\beta_1}$ = 20, $\hat{\beta_2}$ = 0.07, $\hat{\beta_3}$ = 35, $\hat{\beta_4}$ = 0.01, $\hat{\beta_5}$ = -10.

#### (a) Which answer is correct, and why?
For a fixed value of IQ and GPA, males earn more on average than females provided that the GPA is high enough. This is because while the model estimates \$35,000 increase in salary for females, if GPA is higher than 3.5 then the GPA-Gender interaction term for females overwhelms the /$35,000 increase.

#### (b) Predict the salary of a female with IQ of 110 and a GPA of 4.0.
$50 + 20(4.0) + 0.07(110) + 35(1) + 0.01(4.0*110) -10(4.0 * 1) = \$137,100$

#### (c) True or false: Since the coefficient for the GPA/IQ interaction term is very small, there is very little evidence of an interaction effect. Justify your answer
False. "Evidence" for an interaction effect would be evaluated using some test of significance or certainty of the coeffieicent not its numerical scale. The effect may in fact be large because IQ is on a numerically larger scale compared to GPA and the binary 0 or 1 for gender. In addition, IQ is multiplied by GPA, increasing the relative scale of the interaction term.

### 4. I collect a set of data (n = 100 observations) containing a single predictor and a quantitative response. I then fit a linear regression model to the data, as well as a separate cubic regression, i.e. $Y = \beta_0 + \beta_1*X + \beta_2*X^2 + \beta_3*X^3 + \epsilon$.

#### (a)Suppose that the true relationship between X and Y is linear, i.e. $Y = \beta_0 + \beta_1*X + \epsilon$. Consider the training residual sum of squares (RSS) for the linear regression, and also the training RSS for the cubic regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.
This would depend on the training sample and therefore there is not enough information. This is because, while the true relationship between Y and X maybe be linear, the cubic regression may fit the training sample "closer" than the linear regression, therefore giving us a lower training residual sum of squares. Alternatively, the nonlinear curves of the cubic model may overshoot or undershoot the training data.

#### (b) Answer (a) using test rather than training RSS.
The testing resdiual sum of squares would be expected to be lower for the linear model rather than than the cubic regression model. This is because the flexbility of the cubic regression model will to closely track the training data and lead to larger residuals when fit to the testing data.

#### (c) Suppose that the true relationship between X and Y is not linear, but we don’t know how far it is from linear. Consider the training RSS for the linear regression, and also the training RSS for the cubic regression. Would we expect one to be lower than the other, would we expect them to be the same, or is there not enough information to tell? Justify your answer.
There is not enough information to tell. The size of the residuals for each model will be highly dependent on how nonlinear the data is. Small deviations in linearity may not disadvantage the linear regression in comparison to the cubic regression model, however, if the relationship between the predictor and response was in fact quadratic, cubic, or higher, we would expect the cubic model to perform better.

#### (d) Answer (c) using test rather than training RSS.
This answer is similar to part (c). The outcome is highly dependent on the nonlinearity of the population data. One difference is that we would expect the performance of the cubic regression model to perform much better in comparison to the training data (lower test residual sum of squares) as nonlinearity increases.  


### 5. 

### 6.

### 7.

### 8. This question involves the use of simple linear regression on the Auto data set.

#### (a) Use the lm() function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the summary() function to print the results. Comment on the output.
```{r}
#Load data
library(ISLR)
auto = Auto
```

```{r}
#Fit simple linear regression
lm.fit = lm(mpg ~ horsepower, data = auto)
summary(lm.fit)
```
##### (i) Is there a relationship between the predictor and the response?
Given the high F-statistic of the model we can say the model itself is significant. In addition, the p-value of the  horsepower coefficient is lower than 0.01 so we can say the horsepower is a significant predictor and therefore there is a relationship between the predictor and the response.

##### (ii) How strong is the relationship between the predictor and the response?
For every 1 unit increase in horsepower, there is a correspondiing 0.15 decrease in mpg. The model explains 60 percent of the variation in mpg.

##### (iii) The horsepower coefficient is negative (-0.157) and a 99% confidence interval does not include a change of sign, therefore we can say the relationship between horsepower and mpg is negative.

##### (iV) 
```{r}
#mpg prediction for 98 horsepower
mpg_pred_pi = predict(lm.fit, data.frame(horsepower = c(98)), interval = "confidence")
mpg_pred_ci = predict(lm.fit, data.frame(horsepower = c(98)), interval = "prediction")
print(paste0("The point estimate for mpg at 98 horsepower is ",mpg_pred_pi[1], ", while the confidence interval is (", mpg_pred_ci[2], ",", mpg_pred_ci[3], ") and the prediction interval is (",mpg_pred_pi[2], ",",mpg_pred_pi[3], ")."))
```

#### (b)
```{r}
plot(auto$horsepower, auto$mpg)
abline(lm.fit, col="red")
```

#### (C)
```{r}
par(mfrow=c(2,2))
plot(lm.fit)
```
Residuals show positive bias at the lower and upper extremes of mpg, while negative near the median. This could be because the true relationship between horsepower and mpg is polynomial in nature, possibly quadratic. The Normal Q-Q plot demonstrates deviations from normality at the lower and upper extremes of the standardized residuals. Finally there are number of high leverage points but very little outliers.

### 9.This question involves the use of multiple linear regression on the Auto data set.

#### (a) Produce a scatterplot matrix which includes all of the variables in the data set.
```{r}
#Scatterplot Matrix
pairs(auto)
```

#### (b) Compute the matrix of correlations between the variables using the function cor(). You will need to exclude the name variable, cor() which is qualitative.
```{r}
cor(auto[,-which(names(auto) == "name")])
```

#### (c) Use the lm() function to perform a multiple linear regression with mpg as the response and all other variables except name as the predictors. Use the summary() function to print the results.
```{r}
lm.fit2 = lm(mpg ~ . -name, data = auto)
summary(lm.fit2)
```

##### (i) Is there a relationship between the predictors and the response?
The adjusted R-squared of the moel is approximately 0.81 which means the model caputres a 81% of the variation in mpg. The F-statistic is large enough as to reject the null hypothesis that this is no relation between the predictors and the response variable.

##### (ii) Which predictors appear to have a statistically significant relationship to the response?
Displacement, weight, year, and origin are all statistically significant at an alpha level of 0.01.

#### (iii) What does the coefficient for the year variable suggest?
The coefficient for year is 0.75. This suggests that, holding all other predictors fixed, the mpg inceases by 0.75 for each additional year.

###(d) Use the plot() function to produce diagnostic plots of the linear regression fit. Comment on any problems you see with the fit. Do the residual plots suggest any unusually large outliers? Does the leverage plot identify any observations with unusually high leverage?
```{r}
par(mfrow=c(2,2))
plot(lm.fit2)
```
Positive residual outliers exist at the upper end of mpg and at the very lower end. There is a preponderance of negative residual outliers in the middle of the mpd data between 15 and 30. Two obervations are substantially above 0.02 (derived using p+1/n which is a dianostic test for high leverage observations), including observation 14 which has very high leverage.

### (e) Use the * and : symbols to fit linear regression models with interaction effects. Do any interactions appear to be statistically significant?
```{r}
lm.fit3 = lm(mpg ~ . +cylinders:horsepower+ year:origin+weight:acceleration -name, data = auto)
summary(lm.fit3)
```
The cylinders and horsepower interaction effect is significant at an alpha level of .01. Neither the year-origin or weight-accelration is significant at an alpha level of 0.05.


### (f) Try a few different transformations of the variables, such as $log(X)$, $\sqrt{X}$, $X^2$. Comment on your findings.
```{r}
lm.fit4 = lm(mpg ~ (horsepower^2) + log(displacement) +sqrt(year), data = auto )
summary(lm.fit4)
```
Horsepower was raised to a power of two, the log was taken of displacement, and the square root was taken of year. Each of these transformed variables was significant and the model achieve an adjusted R-squared of 0.79 which is close to the full model excluding name.

## 10. This question should be answered using the Carseats data set
```{r}
library(car) #For diagnostics
carseats = Carseats
```

### (a) Fit a multiple regression model to predict Sales using Price, Urban, and US.
```{r}
lm.car = lm(Sales ~ Price + Urban + US, data=carseats)
```

### (b) Provide an interpretation of each coefficient in the model. Be careful—some of the variables in the model are qualitative!
```{r}
summary(lm.car)
```
The model estimates each dollar increase in price results in 0.05 less unit sales. Urban sales locations produce 0.02 less sales while US locations being in an estimated 1.2 more in unit sales.

#### (c) Write out the model in equation form, being careful to handle the qualitative variables properly.

####(d) For which of the predictors can you reject the null hypothesis $H_0 : \beta_j = 0$?
At an alpha level of 0.01 we can reject the null hypothesis for Price and US.

#### (e) On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.
```{r}
lm.car2 = lm(Sales ~ Price + US, data = carseats)
summary(lm.car2)
```

#### (f)How well do the models in (a) and (e) fit the data?
Both models have F-statistics high enough to reject the null hypothesis that the response has no relationship with the predictors. However, both models have low, albeit nearly equal, R-squared values (approximately 0.23). Therefore both models fit the data somewhat poorly.

#### (g) Using the model from (e), obtain 95 % confidence intervals for the coefficient(s).
```{r}
confint(lm.car2)
```

#### (f)  Is there evidence of outliers or high leverage observations in the model from (e)?
```{r}
par(mfrow=c(2,2))
plot(lm.car2)
```

```{r}
outlierTest(lm.car2)
leveragePlots(lm.car2)
influencePlot(lm.car2)
```
The residual plots show approximately normally distributed random residuals. Analysis of the residuals vs. leverage plot show no values with extraordinary Cook's distances. The Bonferroni-adjusted outlier test indicates no outliers, however, it shows the observation with the largest studentized residual is observation 377.

### 11. In this problem we will investigate the t-statistic for the null hypothesis $H_0 : \beta = 0 in simple linear regression without an intercept. To begin, we generate a predictor x and a response y as follows:
```{r}
set.seed(1)
x = rnorm(100)
y = 2*x + rnorm(100)
```

#### (a) Perform a simple linear regression of y onto x, without an intercept. Report the coefficient estimate $\hat{\beta}$, the standard error of this coefficient estimate, and the t-statistic and p-value associated with the null hypothesis $H_0 : \beta = 0. Comment on these results. (You can perform regression without an intercept using the command lm(y ~ x+0).)
```{r}
fit.y_no_intercept = lm(y ~ x + 0)
summary(fit.y_no_intercept)
```
??????

#### (b)  Now perform a simple linear regression of x onto y without an intercept, and report the coefficient estimate, its standard error, and the corresponding t-statistic and p-values associated with the null hypothesis H_0 : \beta = 0. Comment on these results.
```{r}
fit.x_no_intercept = lm(x ~ y + 0)
summary(fit.x_no_intercept)
```
?????

#### (c) What is the relationship between the results obtained in (a) and (b)?


#### (d) For the regression of Y onto X without an intercept, the t-statistic for $H_0 : \beta = 0$ takes the form \hat{\beta}/SE(\hat{\beta}), where $\hat{\beta}$ is given by (3.38), and where $$SE(\hat{\beta}) = \sqrt{\frac{\sum_{i=1}^{n} (y_i - x_i\hate{\beta})^{2}}{(n-1)\sum_{i'}^{2}}}$$. (These formulas are slightly different from those given in Sections 3.1.1 and 3.1.2, since here we are performing regression without an intercept.) Show algebraically, and confirm numerically in R, that the t-statistic can be written as $$\frac{\sqrt{n-1}\sum_{i=1}^{n}x_iy_i}{\sqrt{(\sum_{i=1}^{n}x_{i}^{2})(\sum_{i'=1}^{n}y_{i`}^{2}) - (\sum_{i'=1}^{n}x_{i'}y_{i'})^{2})}}

#### (f) In R, show that when regression is performed with an intercept, the t-statistic for $H_0 : \beta_1 = 0 is the same for the regression of y onto x as it is for the regression of x onto y.

### 12. This problem involves simple linear regression without an intercept.

#### (a) Recall that the coefficient estimate $\hat{{\beta}} for the linear regression of Y onto X without an intercept is given by (3.38). Under what circumstance is the coefficient estimate for the regression of X onto Y the same as the coefficient estimate for the regression of Y onto X?
The coefficient will be the same When the relationship between Y and X is one-to-one and doesn't require a transformation.

#### (b) Generate an example in R with n = 100 observations in which the coefficient estimate for the regression of X onto Y is different from the coefficient estimate for the regression of Y onto X.
```{r}
x = rnorm(100)
y = sqrt(x+10) + rnorm(100)
lm.x_fit_random = lm(y ~ x + 0)
lm.y_fit_random = lm(x ~ y + 0)
summary(lm.x_fit_random)
summary(lm.y_fit_random)
```
#### (c) Generate an example in R with n = 100 observations in which the coefficient estimate for the regression of X onto Y is the same as the coefficient estimate for the regression of Y onto X.
```{r}
x = rnorm(100)
y = x + rnorm(100, mean = 0, sd = 0.5)
lm.x_fit_random = lm(y ~ x + 0)
lm.y_fit_random = lm(x ~ y + 0)
summary(lm.x_fit_random)
summary(lm.y_fit_random)
```
????

### 13. In this exercise you will create some simulated data and will fit simple linear regression models to it. Make sure to use set.seed(1) prior to starting part (a) to ensure consistent results.

#### (a) Using the rnorm() function, create a vector, x, containing 100 observations drawn from a N(0, 1) distribution. This represents a feature, X.
```{r}
set.seed(1)
x = rnorm(100, mean = 0, sd = 1)
```

#### (b) Using the rnorm() function, create a vector, eps, containing 100 observations drawn from a N(0, 0.25) distribution i.e. a normal distribution with mean zero and variance 0.25.
```{r}
eps = rnorm(100, mean = 0, sd = 0.25)
```

#### (c) Using x and eps, generate a vector y according to the model $$Y = -1 + 0.5X + \epsilon$$. What is the length of the vector y? What are the values of \beta_0 and \beta_1 in this linear model?
```{r}
y = -1 + 0.5*x + eps
length(y)
```
The length of the y vector is 100. $\beta_0$ = -1 and $\beta_1 = 0.5$.

#### (d) Create a scatterplot displaying the relationship between x and y. Comment on what you observe.
```{r}
plot(x,y)
```
We observe a positive relationship between x and y. The y-intercept is approximately -1 while one could easily draw a regression line with a slope of 0.5 through the points.

#### (e) Fit a least squares linear model to predict y using x. Comment on the model obtained. How do \hat{\beta_0} and \hat{\beta_1} compare to \beta_0 and \beta_1?
```{r}
fit.random = lm(y ~ x)
summary(fit.random)
```
The intercept is approximately -1.01 and the coefficient of x is approximately 0.499. These are very close to the original programmed values.

#### (f) Display the least squares line on the scatterplot obtained in (d). Draw the population regression line on the plot, in a different color. Use the legend() command to create an appropriate legend.
```{r}
plot(x,y)
abline(fit.random, col = "red")
legend(x = "bottomright", legend = c("Regression Line"), col = "red", lty = 1)
```
#### (g) Now fit a polynomial regression model that predicts y using x and x2. Is there evidence that the quadratic term improves the wmodel fit? Explain your answer.




