---
title: "Lab-Linear-Regression"
author: "Jonathan Bryan"
date: "May 18, 2017"
output: pdf_document
---

3.6.1 Libraries
```{r}
library(MASS)
library(ISLR)
```

3.6.2 Simple Linear Regression
```{r}
boston = Boston
fix(boston)
names(boston)
```


Linear regression using medv as response and lstat as predictor
```{r}
lm.fit = lm(medv ~ lstat, data=boston)
lm.fit

summary(lm.fit) #Run summary statistics on linear model
names(lm.fit) #Determine the names of data stored in lm.fit
coef(lm.fit) #Get the coefficient names and values of lm.fit
confint(lm.fit) #Get 95% confidence intervals for the coefficients

new_data = data.frame(lstat = c(5,10,15))
predict(lm.fit, new_data, interval ="confidence")
predict(lm.fit,new_data,interval = "prediction")
```

Graphical plots
```{r}
plot(boston$medv, boston$lstat)
abline(lm.fit)

#Different plots with adjusted settings
abline(lm.fit, lwd = 3, col ="red")
plot(boston$lstat, boston$medv, col = "red")
plot(boston$lstat, boston$medv, pch = 20)
plot(boston$lstat, boston$medv, pch = "+") #Using plus (+) symbol for scatter plot
plot(1:20, 1:20, pch = 1:20) #Different possible symbols

#Diagnostic plots
par(mfrow= c(2,2))
plot(lm.fit)

plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
plot(hatvalues(lm.fit)) #Leverage diagnostic
which.max(hatvalues(lm.fit)) #Index with largest hat value
```


3.6.3 Multiple Linear Regression
```{r}
#MLR with just lstat and age as predictors
lm.fit = lm(medv ~ lstat + age, data = boston)
summary(lm.fit)

#MLR with all 13 predictors from the Boston dataset
lm.fit_all = lm(medv ~ ., data = boston)
summary(lm.fit_all)
```

Variance inflation factors
```{r}
library(car)
vif(lm.fit_all)
```

Exclude age predictor
```{r}
#MLR excluding age
lm.fit1 = lm(medv ~ . -age, data = boston)
summary(lm.fit1)
```

3.6.4 Interaction Terms
```{r}
lm.fit_interaction = lm(medv ~ lstat*age, data = boston) #The * is equivalent to lstat+age+lstat:age
summary(lm.fit_interaction)
```

3.6.5 Non-linear Transformations of the Predictors
```{r}
lm.fit_non_linear = lm(medv ~ lstat + I(lstat^2), data = boston)
summary(lm.fit_non_linear)
```

Analysis of variance
```{r}
lm.fit = lm(medv ~ lstat, data = boston)
anova(lm.fit,lm.fit_non_linear)
```

Residual analysis
```{r}
par(mfrow=c(2,2))
plot(lm.fit_non_linear)
```

Polynomial transformation
```{r}
lm.fit_polynomial = lm(medv ~ poly(lstat,5), data = boston)
summary(lm.fit_polynomial)
```

Log transformation
```{r}
lm.log = lm(medv ~ log(lstat), data = boston)
summary(lm.log)
```

3.6.6 Qualitative Predictors
```{r}
fix(Carseats)
names(Carseats)
carseats = Carseats
```

```{r}
lm.fit_dummy = lm(Sales ~ . + Income:Advertising + Price:Age, data = Carseats)
summary(lm.fit_dummy)
```

```{r}
attach(Carseats)
contrasts(ShelveLoc)
```

3.6.7 Writing Functions
```{r}
LoadLibraries = function(){
  library(ISLR)
  library(MASS)
  print("The libraries have been loaded.")
}
```

