---
title: 'Lab: Introduction to R'
author: "Jonathan Bryan"
date: "May 7, 2017"
output: html_document
---

Concatenate function c()
```{r}
x <- c(1,3,2,5)
x
```

Concatenate function c()
```{r}
x = c(1,6,2)
x
y = c(1,4,3)
```

Length function length()
```{r}
length(x)
length(y)
x + y #Vector addition
```

List saved objects ls()
```{r}
ls()
```

Delete saved objects rm()
```{r}
rm(x,y)
ls()
rm(list=ls())#Remove all object
```

Create a matrix
```{r}
x = matrix(data=c(1,2,3,4), nrow=2, ncol=2)
x
```

```{r}
x = matrix(data=c(1,2,3,4), 2,2) #Default is to add entries by column
x
```

```{r}
x = matrix(data=c(1,2,3,4), nrow = 2, ncol = 2, byrow=TRUE)
x
```

Important functions
```{r}
sqrt(x) #Square root each entry
x^2 #Take the square of each entry
```

Matrix with normal random variables
```{r}
x = rnorm(n=50)
y = x + rnorm(n=50, mean = 50, sd=.1)
cor(x,y)
```

Using set.seed()
```{r}
set.seed(1303)
q = rnorm(50)
```

Using mean(), var(), and sd()
```{r}
set.seed(3)
y = rnorm(100)
mean(y)
var(y)
sd(y)
```

Using plot()
```{r}
set.seed(1)
x = rnorm(100)
y = rnorm(100)
plot(x,y, xlab="this is the x-axis", ylab="this is the y-axis",main="Plot of X vs Y")
```

Creating sequences of numbers with seq()
```{r}
x = seq(1,10)
x

x = 1:10
x

x = seq(-pi, pi, length=50)
x
```

Creating contour plot with contour()
```{r}
y=x
f = outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
contour(x,y,f, nlevels=45,add=T)
fa = (f-t(f))/2
contour(x,y,fa,nlevels=15)
```

Creating contour plot with image()
```{r}
image(x,y,fa)
persp(x,y,fa)
persp(x,y,fa,theta=30) #Rotation on the z-axis
persp(x,y,fa,theta=30, phi=20) #Rotation on the x-axis
persp(x,y,fa,theta=30, phi=70)
persp(x,y,fa,theta=30, phi=40)
```

Indexing data
```{r}
A = matrix(1:16, nrow=4, ncol= 4)
A
A[2,3] #Select element in row 2 and column 3
A[c(1,3), c(2,4)] #Select rows 1 and 3 as well as columns 2 and 4
A[1:3,2:4] #Select rows 1 through 2 and columns 2 through 4
A[1:2,] #Select rows 1 though 2
A[,1:2] #Select columns 1 through 2
A[1,] #Select row 1
A[-c(1,3),] #Select every row except for 1 and 3
A[-c(1,3), -c(1,3,4)] #Select every row except for 1 and 3 and every column except for 1, 3, and 4
```

Finding the dimensions of a matrix using dim()
```{r}
dim(A)
```

Loading data
```{r}
library(ISLR)
Auto = Auto #Loads ISLR data
fix(Auto) #Generates a spreadsheet of the data for viewing
dim(Auto)
Auto = na.omit(Auto) #Was supposed to go from 397 to 392 but I think these rows have been omitted
dim(Auto)
names(Auto)
```

Creating boxplots plot()
```{r}
plot(Auto$cylinders, Auto$mpg)
#Skip attach() step because this is bad coding
Auto$cylinders = as.factor(Auto$cylinders)
plot(Auto$cylinders, Auto$mpg)
plot(Auto$cylinders, Auto$mpg, col = "red")
plot(Auto$cylinders, Auto$mpg, col ="red", varwidth=T)
plot(Auto$cylinders, Auto$mpg, col = "red", varwidth=T, xlab = "Cylinders", ylab="MPG")
```

Creating histograms with hist()
```{r}
hist(Auto$mpg)
hist(Auto$mpg, col=2) # col=2 has same effect as col="red"
hist(Auto$mpg, col=2, breaks = 15) 
```

Creating scatterplot matricies with pairs()
```{r}
pairs(Auto)
pairs(~ mpg + displacement + horsepower + weight + acceleration, Auto) #Do a scatterplot on a subset of variables
```

Using identify()
```{r}
plot(Auto$horsepower, Auto$mpg)
identify(Auto$horsepower, Auto$mpg, Auto$name)
```

Using summary()
```{r}
summary(Auto)
summary(Auto$mpg)
```

